Follow-up Samples

### 2017-04-19 submission/manuscript.tex
small extra corrections to discussion

The microbiome changes associated with carcinoma were found to be
towards those with normal colons. This would suggest that treatment for
CRC is not only successful for removing the carcinoma but also at
removing bacteria associated with them. This is important due to the
numerous studies that have pointed towards a link between the community
and tumorgenesis {[}10--12{]}. The importance of the community to this
process is further highlighted in our data {[}Figure 2, 4, \& S3{]}.
Interestingly, those who had an increase in their positive probability
of carcinoma had no change in their median value for oral pathogens
found within the carcinoma model but had a decrease in non-oral
pathogens {[}Figure S4{]}. Although this finding was not statistically
significant, this trend would further highlight the importance of
community to tumorgenesis. Future studies with larger power may be able
to confirm these reported observations.

Although we were not able to identify specific OTUs due to the
heterogeneity of the population. This is a common problem that occurs in
other fields of microbiome study {[}28{]}. This made it very difficult
to identify individual OTUs associated with treatment and recovery
towards a normal colon. To overcome this large individual heterogeneity
we used RF models that made it easier to identify consortia of
microbiota that were associated with treatment and recovery. A potential
future direction would be to study whether specific gene signatures
associated with these consortia of bacteria have a role to play in the
recurrence of CRC.

Despite some of these stated shortcomings our findings add to the
existing scientific knowledge on CRC and the microbiome: That there is a
measurable change in the bacterial community after adenoma, advanced
adenoma, or carcinoma treatment. Further, individuals with carcinoma had
a larger change than adenoma or advanced adenoma indivdiuals and this
change in community was towards those with normal colons. Our data
provides additional evidence on the importance of the community on
recovery of the microbiome after CRC treatment and opens interesting
avenues of research into how these changes may affect recurrence.

>>>>

The microbiome changes associated with carcinoma were found to be
towards those with normal colons. This would suggest that treatment for
CRC is not only successful for removing the carcinoma but also at
removing bacteria associated with them. This is important due to the
numerous studies that have pointed towards a link between the community
and tumorgenesis {[}10--12{]}. The importance of the community to this
process is further highlighted in our data {[}Figure 2, 4, \& S3{]}.
Interestingly, those who had an increase in their positive probability
of carcinoma had no change in their median value for oral pathogens
within the carcinoma model but had a decrease in bacteria that were
higher in carcinoma {[}Figure S4{]}. Although this finding was not
statistically significant, this trend further highlights the importance
of community to tumorgenesis. Future studies with larger power may be
able to confirm these reported observations.

Although we were not able to identify specific OTUs due to the
heterogeneity of the population. This is a common problem that occurs in
other fields of microbiome study {[}28{]}. This makes it very difficult
to identify individual OTUs associated with treatment and recovery
towards a normal colon. To overcome this large individual heterogeneity
we used RF models that made it easier to identify consortia of
microbiota that were associated with treatment and recovery. One
potential future direction would be to study whether specific gene
signatures associated with these consortia of bacteria are directly
involved with the pathogenesis of CRC recurrence.

Despite some of these stated shortcomings our findings add to the
existing scientific knowledge on CRC and the microbiome: That there is a
measurable change in the bacterial community after adenoma, advanced
adenoma, or carcinoma treatment. Individuals with carcinoma had a larger
change than adenoma or advanced adenoma individuals and this change in
community was towards those with normal colons. Our data provides
additional evidence on the importance of the community on recovery of
the microbiome after CRC treatment and opens interesting avenues of
research into how these changes may affect recurrence.
